---
title: "ML project"
output: html_document
---

###Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

###Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

###What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

###Loading data
```{r load data,cache=T,include=T}
setwd("~/R/Assignment")
library(dplyr)
library(caret)
training<- read.csv('pml-training.csv',stringsAsFactors = F)%>%tbl_df()
testing<- read.csv('pml-testing.csv',stringsAsFactors = F)
dimension<- c('observations'=nrow(training),'variables'=ncol(training))
print (dimension)
```

####There are `r nrow(training)` observations of the training data set, and there are `r ncol(training)` variables in the training data set. Apprently, we cannot include all the variables in the mechine learning algorithm. 

###Cleaning data
```{r clean data, cache=TRUE}
y<- vector()
for ( i in 1:ncol(training)){
  x<- sum(is.na(training[,i]))
  if (x/nrow(training)>0.7){
    y<- c(y,F)
  } else {
    y<- c(y,T)
  }
}
training_clean<- training[,y]
training_clean<- training_clean[,sapply(training_clean,function(x) !is.character(x))]%>%mutate(class=as.factor(training$classe))
training_clean<- training_clean[,-c(1:4)]
```

####First I get rid of variables that have more than 70% NA rate. Then I get rid of variables whose class are character. Finally, because the first 4 columns are not related to actual movement mearsurements, I aslo get rid of them.

###Building Machine Learning Algorithm
```{r ML, cache=T}
library(randomForest)
train_partition<- createDataPartition(y = training_clean$class,p=0.7,list=F)
subTrain<- training_clean[train_partition,]  #subTrain group for training model
subTest<- training_clean[-train_partition,]  # subTest group for cross-validation
rf_model<- randomForest(class~.,data = subTrain) # randomForest method is used for building model
confusionMatrix(predict(rf_model,newdata = subTest),subTest$class)
```

####The training data is divided into two groups, subTrain and subTest. 70% of the training group is randomly assigned to subTrain which is used for training the model. And the rest 30% is assigned to subTest which is used for cross-validation to test the accuracy of the model. I choose the Random Forest method for model building. After cross-validation, it turns out that the accuracy of the random forest method is very high. The out-of-sample error for cross-validation set is **0.0054%**

### Prediction with the model
```{r prediction, cache=T}
predict(rf_model,newdata = testing)
```


